The deployment of Reinforcement Learning (RL) in psychiatric digital interventions faces an inherent ``Adaptivity Trap'': highly adaptive agents optimize for proximal rewards by converging on niche, deterministic policies. In live trials, this behavior eliminates the stochasticity necessary for Off-Policy Evaluation (OPE), fundamentally violating algorithm fidelity and degrading the scientific utility of the trial. To resolve the tension between optimization and evaluation, we introduce a Dynamic Fidelity-Constrained RL architecture. Our framework utilizes Neural Stochastic Differential Equations (Neural SDEs) to model non-stationary clinical volatility on compact state-spaces, while simultaneously capturing informative missingness via a coupled Hawkes-modulated observation process. A Dynamic Fidelity Controller continually estimates real-time statistical power---using Marginalized Importance Sampling to approximate Effective Sample Size (ESS)---and dynamically modulates the agent’s exploration bounds. During stable periods, the policy expands to collect high-variance causal data; during high-volatility states or crisis-indicating non-responses, it strictly contracts to a pre-defined baseline protocol via Reverse KL divergence penalties. Evaluation via simulated synthetic cohorts---stress-tested using Bayesian non-parametric anomalies (HDP-HMM)---demonstrates that our dynamically constrained agent successfully preserves ESS across trial horizons. This preservation enables robust, unbiased Offline OPE (using Marginalized Doubly Robust estimators), thereby ensuring both participant safety and the preservation of rigorous scientific validity without fatally compromising longitudinal clinical outcomes.
